### 2017-07-08 sheets/talk-10/sheet-10.tex
Extended solutions.

\begin{exercise}
  Recall the following re-formulation of the Yoneda principle: For any locally small category $\cat A$ and objects $A,A'\in\cat A$:
  \begin{align*}
    \cat A(B,A)\iso \cat  A(B,A')\text{ naturally in } B\in\cat A
  \end{align*}
implies $A\iso A'$. Show that naturality is really necessary.
\hint{There is a counterexample with two objects.}
\end{exercise}

@@

\begin{exercise}
  Let $\cat B$ be a category and $J\from\cat C\to D$ a functor. There is an induced functor
  \begin{align*}
    J\of -\from[\cat B,\cat C]\to [\cat B,\cat D]
  \end{align*}
  defined by composition with $J$.
  \begin{enumerate}
  \item Show that if $J$ is full and faithful then so is $J\of -$.
  \item Deduce that if $J$ is full and faithful and $G,G'\from\cat B\to\cat C$ with $J\of G\iso J\of G'$ then $G\iso G'$.
  \item Now deduce that right adjoints are unique: if $F\from \cat A\to\cat B$ and $G,G'\from\cat B\to \cat A$ with $F\dashv G$ and $F\dashv G'$ then $G\iso G'$. 
    \hint{The Yoneda embedding is full and faithful.}
  \end{enumerate}
\end{exercise}

@@

>>>>

\begin{exercise}
  Recall the following re-formulation of the Yoneda principle: For any locally small category $\cat A$ and objects $A,A'\in\cat A$:
  \begin{align*}
    \cat A(B,A)\iso \cat  A(B,A')\text{ naturally in } B\in\cat A
  \end{align*}
implies $A\iso A'$. Show that naturality is really necessary.
\hint{There is a counterexample with two objects.}
\end{exercise}
\begin{answer}
consider the following category $\cat A$: 
\begin{align*}
  \xymatrix{
  A \ar@(ul,ur)^{1_A,f} \ar@/^0.3pc/@[][r]^{f_1,f_2} & B \ar@/^0.3pc/@[][l]^{g_1,g_2} \ar@(ul,ur)^{1_B,g}
                                                       } 
  \end{align*}
  i.e. between any pair of object there are two arrows, where the composition of $f_i$ and $g_j$ ($i,j\in\{1,2\}$) is given by
  \begin{align*}
    f_i\of g_j&=g
  \end{align*}
  and the remaining compositions can be chosen arbitrarily.

  Obviously,
  \begin{align*}
    \cat A(C,A)\iso\cat A(C,B)
  \end{align*}
  for all $C\in\cat A$ because all homsets contain exactly two elements. However $A\not\iso B$ because by construction neither $g_1$ nor $g_2$ (which are the only maps between $B$ and $A$) have a left inverse.
\end{answer}

@@

\begin{exercise}
  Let $\cat B$ be a category and $J\from\cat C\to\cat D$ a functor. There is an induced functor
  \begin{align*}
    J\of -\from[\cat B,\cat C]\to [\cat B,\cat D]
  \end{align*}
  defined by composition with $J$.
  \begin{enumerate}
  \item Show that if $J$ is full and faithful then so is $J\of -$.
  \item Deduce that if $J$ is full and faithful and $G,G'\from\cat B\to\cat C$ with $J\of G\iso J\of G'$ then $G\iso G'$.
  \item Now deduce that right adjoints are unique: if $F\from \cat A\to\cat B$ and $G,G'\from\cat B\to \cat A$ with $F\dashv G$ and $F\dashv G'$ then $G\iso G'$. 
    \hint{The Yoneda embedding is full and faithful.}
  \end{enumerate}
\end{exercise}
\begin{answer}
  \begin{enumerate}
  \item Clearly, $J\of-$ is defined on arrows $\alpha\from F\to G$ for $F,G\from\cat B\to\cat C$ by
    \begin{align*}
      J\of -(\alpha)&=\alpha'
    \end{align*} 
    where $\alpha$ is a natural transformation $J\of F\to J\of G$ defined by
    \begin{align*}
      \alpha'_B=J(\alpha_B)\from J\of F(B)\to J\of G(B)
    \end{align*}
    for $B\in\cat B$.

    Full and faithfulness of $J$ means that for each $C,C'\in\cat C$ the function
    \begin{align*}
      J_{C,C'}\from\cat C(C,C)\to \cat D(J(C),J(C'))\\
      f&\mapsto J(f)
    \end{align*}
    is bijective.

    To show that $J\of-$ is full and faithful, let $F,G\from\cat B\to\cat C$ be natural transformations and $\beta\from J\of F\to J\of G$ be a natural transformation.
    Define, a natural transformation $\alpha\from F\to G$ by 
    \begin{align*}
      \alpha_B=J_{F(A),G(A)}^{-1}(\beta_B)
    \end{align*}
    for $B\in\cat B$. This is well-defined by full- and faithfulness and defines indeed a natural transformation.

    Clearly, $J\of -(\alpha)=\beta$. This shows fullness of $J\of -$. For faithfulness assume a natural transformation $\alpha'\from F\to G$ and suppose 
    \begin{align*}
      J\of-(\alpha)=J\of-(\alpha')
    \end{align*}
    By definition this means
    \begin{align*}
      J(\alpha_B)=J(\alpha'_B)
    \end{align*}
    for any $B\in\cat B$. Faithfulness of $J$ implies $\alpha_B=\alpha'_B$ for any $B\in\cat B$, i.e. $\alpha=\alpha'$.
  \item \label{ex:lem} $J\of -\from[\cat B,\cat C]\to [\cat B,\cat D]$ is full and faithful by the previous part. Hence for $G,G'\in [\cat B,\cat C]$, $G\iso G'$ if and only if $J\of G=J\of-(G)\iso J\of-(G')=J\of G'$.
  \item By adjointness we get
    \begin{align*}
      H_{G(B)}(A)=\cat A(A,G(B))\iso\cat B(F(A),B)\iso\cat A(A,G'(B))=H_{G'(B)}(A)
    \end{align*}
    naturally in $A\in\cat A$ and $B\in\cat B$. In particular,
    \begin{align*}
      (H_-\of G)(B)=H_{G(B)}\iso H_{G'(B)}=(H_-\of G')(B)
    \end{align*}
    naturally in $B$, i.e. $H_-\of G\iso H_-\of G'$. Due to the fact that the Yoneda embedding $H_-$ is full and faithful and part \ref{ex:lem}, $G\iso G'$.

    % By adjointness we get
%     \begin{align*}
%       \cat A(A,G(B))\iso\cat B(F(A),B)\iso\cat A(A,G'(B))
%     \end{align*}
%     naturally in $A\in\cat A$ and $B\in\cat B$. % This means 
%     % \begin{align*}
%     %   H_{G(B)}(A)\iso H_{G'(B)}(A)
%     % \end{align*}
%     % naturally in $A$, i.e. $(H_-\of G)(B)=H_{G(B)}\iso H_{G'(B)}=(H_-\of G')(B)$.
% This means there is a natural transformation $\eta_B$ that is an isomorphism and for which
%     \begin{align*}
%                 H_{G(B)}(A)\xrightarrow{\eta_{B,A}}H_{G'(B)}(A)
%     \end{align*}
%     naturally in $A$ and $B$.

%     As a consequence $H_-(G(B))=H_{G(B)}\iso H_{G'(B)}=H_-(G'(B))$ and the isomorphism is given by $\eta_B$. By the Yoneda principle, $G(B)\iso G'(B)$.

%     This isomorphism is even natural in $B$: Since $\cat A(A,G(B))\iso\cat A(A,G'(B))$ naturally in $B$ the following commutes for any $f\from B\to B'$:
%     \begin{align*}
%       \xymatrix{
%       H_{G(B)} \ar[r]^{H_{G(f)}}\ar[d]^{\eta_B} & H_{G(B')} \ar[d]_{\eta_{B'}}\\
%       H_{G(B')} \ar[r]^{H_{G'(f)}} & H_{G'(B')} 
%                                         } 
%   \end{align*}
%   Because the Yoneda embedding $H_-$ is full and faithful this means
%   \begin{align*}
%     \xymatrix{
%     G(B) \ar[r]^{G(f)}\ar[d]^{\zeta_B} & G(B') \ar[d]_{\zeta_{B'}}\\
%     G(B') \ar[r]^{G'(f)} & G'(B') 
%                                    } 
%   \end{align*}
%   commutes where $\zeta_B$ and $\zeta_{B'}$ are the isomorphism corresponding to $\eta_B$ and $\eta_{B'}$ respectively under $H_-$. Another consequence is that $G(B)\iso G'(B)$ naturally in $B$, i.e. $$.
  \end{enumerate}
\end{answer}

@@

\begin{answer}
  Let $F,G,\alpha,\beta$ be given as stated above. By the naturality statement in the Yoneda lemma we have that
  \begin{align*}
    \xymatrix{
    [\cat A^\op,\Set](H_A,F) \ar[r]^{\alpha\of -} & [\cat A^\op,\Set](H_A,G) \\
    F(A) \ar[r]^{\alpha_A} \ar[u]^{\iso} & G(A) \ar[u]^{\iso}
                             } 
  \end{align*}
commutes for any $A\in\cat A$. In particular, with the definitions from the proof of the Yoneda lemma we have that 
\begin{align*}
    \xymatrix{
    [\cat A^\op,\Set](H_A,F) \ar[r]^{\alpha\of -} & [\cat A^\op,\Set](H_A,G) \ar[d]^{\widehat{-}} \\
    F(A) \ar[r]^{\alpha_A} \ar[u]^{\widetilde{-}} & G(A) 
                             } 
\end{align*}
commutes for any $A\in\cat A$. Consequently, for any $A\in\cat A$ and $a\in F(A)$,
\begin{align*}
  \alpha_A(a)=\widehat{\alpha\of\widetilde a}.
\end{align*}
Analogously,
\begin{align*}
  \beta_A(a)=\widehat{\beta\of\widetilde a}.
\end{align*}
By assumption ($\widetilde a\from H_A\to F$)
\begin{align*}
  \alpha\of\widetilde a=\beta\of\widetilde a
\end{align*}
holds and therefore
\begin{align*}
  \alpha_A(a)=\beta_A(a)
\end{align*}
for any $A\in\cat A$ and $a\in F(A)$, which means $\alpha=\beta$.
\end{answer}
